\documentclass{book}
\author{Jonathan Pearson}
\title{Sorm Developers' Manual}
\begin{document}

\tableofcontents

\chapter{Overview}
Sorm is a combination code generator and library to provide a straightforward,
transparent mapping between a relational database and Java code. It provides a
similar function as Hibernate or myBatis, but aims for a more direct,
transparent interface.

\section{Introduction}
I created Sorm because I was dissatisfied with existing ORM layers. Hibernate
was too complex and thick, and felt like it was trying to make the database
invisible. It has its own SQL dialect and uses code weaving, with I distrust.
I also tried myBatis, but could not get the code generator to work.

With Sorm I wanted to make configuration and usage easy. I discovered while
writing the code generator that it is virtually impossible to make the
configuration easy (there are simply too many necessary options), but I still
tried to make it as simple as possible. The code generator is likewise simple,
acting as a simple compiler with only input/output file options (and working
with stdin/stdout by default).

\section{Overview}
Sorm is in two pieces, a code generator and a library.

The code generator is used during project setup/compilation to create Java
source files representing the objects of your data model that you want to be
persisted in a database.

The library is used by your code and the generated code to manage database
connections and an optional object cache.


\chapter{Configuration}
Configuration is through XML files, one per data object that you want to define.

\section{Input File Reference}
This is the layout of the file, with all available options described:

Top-level: sorm
\begin{enumerate}
    \item @pkg [String, Required] The name of the Java package that will contain
          the generated class. Use dot (.) as the separator character.
    \item @accessor [String, Optional, Default ``public''] The accessor for the
          class (public or empty).
    \item @name [String, Required] The name of the generated class.
    \item @orm-accessor [String, Optional, Default ``public''] The accessor for
          the generated ``Orm'' static inner-class.
    \item @super [String, Optional, Default null] The full name of the
          superclass of the generated class, including the package. Use dot (.)
          as the separator character.
    \item field[1..*] Describe the fields of the generated class.
    \begin{enumerate}
        \item @accessor [String, Optional, Default ``private''] The accessor for
              the field (public, protected, empty for default, or private).
        \item @type [String, Required] The Java type of the field. If using a
              class outside of java.lang, make sure to provide the full name,
              including the package.
        \item @name [String, Required] The name of the field.
        \item @primary [Boolean, Optional, Default ``false''] Whether this is
              the primary identifier for the object in the database. Exactly one
              primary field is required, and only one is supported.
        \item @sql-type [String, Optional, Default @type] The type of the field.
              Must be the name of one of the enum constants in
              net.jonp.sorm.codegen.SQLType.
        \item @sql-column [String, Optional, Default @name] The name of the
              column in the database.
        \item @nullable [Boolean, Optional, Default false] Whether the
              field/column accepts null values.
        \item @from-super [Boolean, Optional, Default false] Whether the field
              actually exists on the superclass and should not be part of the
              generated class. If true, no field declaration or accessors will
              be emitted, but the field will be accessible through SQL
              statements.
        \item get[0..1] Describes both the getter for this field, and the way to
              convert from the field type to the SQL type.
        \begin{enumerate}
            \item @name [String, Optional, Default ``get[Field name]''] The name
                  for the getter. The default capitalizes the first letter of
                  the field name and prefixes it with ``get''.
            \item @accessor [String, Optional, Default ``public''] The accessor
                  of the getter function.
            \item @override [Boolean, Optional, Default ``false''] Whether to
                  include an ``@Override'' annotation on the generated getter
                  function.
            \item @super [Boolean, Optional, Default ``false''] Whether this
                  getter is actually implemented on the superclass. If true,
                  no getter function is emmitted for the generated class.
            \item (Text) [String, Optional, Default ``\%\{\}.get[Field name]()'']
                  An optional conversion from the field getter to the SQL type
                  for the column. The object being accessed may be referenced
                  as ``\%\{\}''. For example, if your field is a java.util.Date,
                  but your SQL column is a java.sql.Timestamp, your getter text
                  should probably be something like
                  ``new java.sql.Timestamp(\%\{\}.getDate().getTime())''.
        \end{enumerate}
        \item set[0..1] Describes both the setter for this field, and the way to
              convert from the SQL type to the field type.
        \begin{enumerate}
            \item @name [String, Optional, Default ``set[Field name]''] The name
                  for the setter. The default capitalizes the first letter of
                  the field name and prefixes it with ``set''.
            \item @accessor [String, Optional, Default ``public''] The accessor
                  of the setter function.
            \item @override [Boolean, Optional, Default ``false''] Whether to
                  include an ``@Override'' annotation on the generated setter
                  function.
            \item @super [Boolean, Optional, Default ``false''] Whether this
                  setter is actually implemented on the superclass. If true,
                  no setter function is emmitted for the generated class.
            \item (Text) [String, Optional, Default
                  ``\%\{\}.set[Field name](\%\{[Field name]\})''] An optional
                  conversion from the SQL type for the column to the type
                  accepted by the setter. The object being accessed may be
                  referenced as ``\%\{\}'', and fields from the SQL may be
                  referenced as ``\%\{fieldName\}''. For example, if your field
                  is a java.util.Date, but your SQL column is a
                  java.sql.Timestamp, your setter text should probably be
                  something like
                  ``\%\{\}.setDate(new java.util.Date(\%\{date\}.getTime()))''.
        \end{enumerate}
        \item link[0..1] If this is a linked/mapped type, describes the link.
        \begin{enumerate}
            \item @mode [String, Required] The type of link. One of ``None'',
                  ``OneToOne'', ``OneToMany'', ``ManyToOne'', or ``ManyToMany''.
            \begin{itemize}
                \item None means this is not a linked type. There is no
                reason to specify a link element with a link type of None
                for a field.
                \item OneToOne means that this object contains a reference
                to another object, and that the other object contains a
                reference to this object. The field must be nullable in
                this case (otherwise how would you construct the pair? One
                of them must be created before the other, and the field
                would need to be null in the time between).
                \item OneToMany means that this object contains a
                collection of the linked objects. This is generally
                implemented by referencing this object in the table
                containing the other objects. A collection subelement
                must be present in this case.
                \item ManyToOne describes the opposite side of a OneToMany
                linkage, but otherwise acts like a OneToOne relationship;
                this object contains a single reference to another type.
                \item ManyToMany describes a type of mapping in which a
                separate table is used to pair objects of two types,
                with one row (or sometimes two) used to designate each
                mapping. In the two row case, each of the two rows has
                the same pair in the opposite order, when the two objects
                being mapped are of the same type. A collection subelement must
                be present in this case. Fields with this type of linkage will
                emit two additional Orm functions named ``map[Fieldname]'' and
                ``unmap[Fieldname]]'', which create and remove mappings,
                respectively.
            \end{itemize}
            \item @key-type [String, Required] The Java type of the key used to
                  match objects of the linked type. Likely the type of the
                  primary key of the object.
            \item @type [String, Required] The Java type of the linked object.
            \item @sql-type [String, Optional, Default @key-type] The SQL type
                  of the key. The name of an enum from
                  net.jonp.sorm.codegen.SQLType.
            \item collection[0..1] Describes the linked collection for *ToMany
                  linkages.
            \begin{enumerate}
                \item read[1..1] Describes how to read the collection of linked
                      objects.
                \begin{itemize}
                    \item r[1..*] The SQL query used to read the collection for
                          a specific SQL dialect.
                    \begin{itemize}
                        \item @dialect [String, Optional, Default ``*''] The
                              name of the dialect for this query.
                        \item (Text) The SQL query used to read the collection.
                              Must return at least a single column named ``id'',
                              although any additional columns will be ignored.
                              Fields may be referenced as ``\%\{fieldName\}''.
                    \end{itemize}
                \end{itemize}
                \item create[0..1] Only necessary for ManyToMany linkages.
                      Describes how to add a new mapping to the database.
                \begin{itemize}
                    \item c[1..*] Describes how to add a new mapping for a
                          specific SQL dialect.
                    \begin{itemize}
                        \item @dialect [String, Optional, Default ``*''] The
                              name of the dialect for this statement.
                        \item (Text) The SQL query used to create a new linkage.
                              Fields on this object may be referenced as
                              ``\%\{1.fieldName\}'', and fields on the object
                              being linked to this object may be referenced as
                              ``\%\{2.getFunction():[nullable ]type\}'', where
                              the type is the name of one of the enum constants
                              in net.jonp.sorm.codegen.SQLType, and the presense
                              of the ``nullable'' keyword indicates that the
                              getter may return null and the column may be null.
                              For example, ``\%\{2.getId():Integer\}'' will use
                              the Integer function ``getId()'' on the linked
                              object when building the statement.
                    \end{itemize}
                \end{itemize}
                \item delete[0..1] Only necessary for ManyToMany linkages.
                      Describes how to remove a mapping from the database.
                \begin{itemize}
                    \item d[1..*] Describes how to remove a mapping for a
                          specific SQL dialect.
                    \begin{itemize}
                        \item @dialect [String, Optional, Default ``*''] The
                              name of the dialect for this statement.
                        \item (Text) The SQL query used to remove the linkage.
                              Fields on this object may be referenced as
                              ``\%\{1.fieldName\}'', and fields on the linked
                              object may be referenced as
                              ``\%\{2.getFunction():[nullable ]type\}''. See
                              the create element above for how this works.
                    \end{itemize}
                \end{itemize}
            \end{enumerate}
        \end{enumerate}
    \end{enumerate}
    \item create[0..1] Describes how to insert a new one of these objects into
          the database.
    \begin{enumerate}
        \item c[1..*] Describes the insert command for a specific SQL dialect.
        \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
                  dialect for this statement.
            \item (Text) The SQL query used to insert the object into the
                  database. Fields may be referenced as ``\%\{fieldName\}'', but
                  remember that the ID has not been set yet.
        \end{enumerate}
    \end{enumerate}
    \item pk[0..1] Describes how to retrieve the ID of the last object inserted
          into the database.
    \begin{enumerate}
        \item pk[1..*] Describes the query for a specific SQL dialect.
        \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
                  dialect for this statement.
            \item (Text) The SQL query used to retrieve the last inserted ID.
                  Must return at least one column named ``id''; other columns
                  will be ignored. Fields may be referenced as
                  ``\%\{fieldName\}'', but remember that the ID has not been set
                  yet.
        \end{enumerate}
    \end{enumerate}
    \item read[0..1] Describes how to read an object out of the database, given
          its ID.
    \begin{enumerate}
        \item r[1..*] Describes the read query for a specific SQL dialect.
            \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
                  dialect for this query.
            \item (Text) The SQL query used to retrieve the object. Must return
                  one column for each field defined for the object, named by the
                  ``sql-column'' parameter for that field. The key value may be
                  referenced by ``\%\{\}'' in the SQL query.
        \end{enumerate}
    \end{enumerate}
    \item update[0..1] Describes how to update an object in the database.
    \begin{enumerate}
        \item u[1..*] Describes the query for a specific SQL dialect.
        \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
                  dialect for this statement.
            \item (Text) The SQL statement used to update the given object.
                  Fields may be referenced as ``\%\{fieldName\}''.
        \end{enumerate}
    \end{enumerate}
    \item delete[0..1] Describe how to delete an object in the database.
    \begin{enumerate}
        \item d[1..*] Describes how to delete an object for a specific dialect.
        \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
              dialect for this query.
            \item (Text) The SQL statement used to delete the given object.
                  Fields may be referenced as ``\%\{fieldName\}''.
        \end{enumerate}
    \end{enumerate}
    \item query[0..*] Describes named queries on the database. These each emit
          a single function which reads keys for the generated object out of
          the database.
    \begin{enumerate}
        \item @accessor [String, Optional, Default ``public''] The accessor for
              the function that implements this query.
        \item @name The name of the getter function that implements this query.
        \item param[0..*] Descriptions of the parameters of the function that
              implements this query.
        \begin{enumerate}
            \item @type The Java type of this parameter.
            \item @name The name of this parameter.
            \item @sql-type The SQL type of this parameter.
            \item set[0..1] Describes how to convert the Java parameter to the
                  SQL value used in the query.
            \begin{enumerate}
                \item (Text) How to convert the Java parameter to the SQL value.
                      The parameter may be referenced as ``\%\{\}''. For
                      example, an input parameter of java.util.Date could be
                      converted to a SQL value of java.sql.Timestamp with
                      ``new java.sql.Timestamp(\%\{\}.getTime())''.
            \end{enumerate}
        \end{enumerate}
        \item q[1..*] Describes how to query value for the specific dialect.
        \begin{enumerate}
            \item @dialect [String, Optional, Default ``*''] The name of the
                  dialect for this query.
            \item (Text) The query to retrieve keys matching the parameters.
                  Parameters may be referenced as ``\%\{parameterName\}''. Must
                  return at least a column named ``id''. Other returned will be
                  ignored.
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

\end{document}

